---
title: "Lab 8 - Linear Regression"
author: "Becky Clow"
date: "11/25/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Linear Regression
Today we will
(a) Simplify data for easier data analysis
(b) Create an exploratory scatterplot
(c) Perform linear regression
(d) Evaluate model diagnostics
(e) Explore model fit and statistics
(f) Prepare a final graph
(g) Use the model output to make predictions
(h) Find Pearson’s r (correlation)

```{r}

library(tidyverse)
salmon <- read_csv("salmon_imports.csv")

```

##Exploratory scatterplot 
To look at the relationship and ensure it is linear so we can do regression

```{r}

salmon_scatter <- ggplot(salmon, aes(x = year, y = salmon_fresh)) +
  geom_point()
salmon_scatter

#Overall, does it look like describing the relationship between year and fresh salmon imports using linear regression makes sense? Do you have any concerns about using linear regression to describe the relationship?

#over all it is linear but there are a few jogs in the data that make linear regression maybe in appropriate.

```


##Linear Regression on the Salmon data 


```{r}
#Is there a linear relationship between volume of salmon imported and year

salmon_model <- lm(salmon_fresh ~ year, data = salmon)
salmon_model #shows intercept and slope

#model equation
# Imports(tons) = -14982940 + 7550(year)


```
What does the slope mean in the context of this model?
The slope is the volume(tons) change of salmon imported to the US per year 

What does the y-intercept mean in the context of this model? Why is that concerning? What does thismean about extrapolating this model for past values?
The y-intercept is the volume of salmon at year zero but it is a negative number which is not possible in this context. We can't use this model to extrapolate to the past because of this negative value.


##Model diagnostics
```{r}

plot(salmon_model) #shows diagnostic plots that describe more about the model

#Do residuals appear normally distributed?
#they are not completely normally distributed,they spread out towards the end

#Any concerns about heteroscedasticity or outliers?
#They are not spread evenly, the outliers increase as you continue down

```

##Explore model fit and significance
```{r}
summary(salmon_model)

#Does year significantly predict salmon imports?
#yes the lower p-value of <2.2e-16 is sig

#What does the R2 value actually mean in words?
#0.9478 means that the salmon volume is 94% explained by the year

#What proportion of the variance in salmon imports is NOT explained by year?
#about 0.06

```


##Prepare final graph

```{r}

salmon_final_graph <- ggplot(salmon, aes(x = year, y = salmon_fresh)) +
  geom_point() +
  geom_smooth(method = lm, se = TRUE, size = 0.5, color = "gray20") + #adds the linear model with a standard error
  theme_bw() +
  scale_x_continuous(expand = c(0,0), limits = c(1990, 2017)) +
  labs(x = "Year", y = "U.S. Annual Fresh Salmon Imports\n(1,000 pounds)")

salmon_final_graph

```

##Make Predictions using the model
Let’s say we want to use our model to make predictions for U.S. fresh salmon imports in years 2022, 2024, 2026 and 2028.

• Option 1 (not recommended): Plug each of those years into the model equation separately to calculate
the expected U.S. imports.
• Option 2 (recommended): Create a data frame containing the ‘new values’ that you want to make
predictions for, feed that into the predict() function

```{r}

new_years <- data.frame(year = c(2022, 2024, 2026, 2028)) #creating a new data frame with just one colum with the given years

#to find values for the years you want use predict()
future_predict <- predict(salmon_model, newdata = new_years, interval = "confidence") #cofidence for the 95% CI, reports the upper and lower bounds
future_predict

#creating a dataframe with the predicted values and the new_years dataframe
predictions <- data.frame(new_years, future_predict)
predictions

```

##Find Pearson's r (correlation)
Use the cor.test() function to find Pearson’s r for the linear relationship described between year and salmon
imports.

```{r}

salmon_r <- cor.test(salmon$year, salmon$salmon_fresh)
salmon_r

#Would you describe this as a weak/strong negative/positive correlation?
#strong positive correlation

```

##Write a concluding statement about the relationship

